---
title: "Data Exploration"
format: html
editor: source
---

```{r setup}
library(dplyr)
library(lubridate)

dta <- read.csv(file = "~/ERM/data_clean/logs_with_devices.csv")
```

```{r}
dta |> head() |> View()
```

```{r}
usage <- dta |> 
  filter(!is.na(point_id)) |> 
  mutate(
    timestamp = dmy_hms(time_raw),
    date = as_date(timestamp),
    hour = hour(timestamp),
    weekday = wday(timestamp, label = TRUE, week_start = 1),
    month = month(timestamp),
    visitor_id = card_hex
  )

```


# TODO: 

- [] go through the research questions and plan how to answer them 
- [] how to map the data points onto the floor plan
    - device_id 
    - a dynamic floorplan
- [] connect metadata to logs 
- [] parquet file + pipeline automation + cleanup 
- [X] network data - model the individual 
- [] network data - compute all paths for all visitors 

```{r}
usage |> tail()
```

# number of visitors

```{r}
usage |> count(visitor_id) |> #93,595 
  filter(n != 1) |>  # 86,264 
  arrange(desc(n))
```


```
visitor_id          n 
03	                28812			
04	                13967			
02	                11048			
06	                8958			
07	                6134			
05	                2862			
032020203247523845	754			
04202020333447394c	750	
```
This is a bit strange -> can a person realistically have this many entries? (differentiate by date?) 
Also what does the shorter entry mean ("03" vs. "032020203247523845")? 

```{r}
usage |> count(visitor_id, date) |> #93,595 
  filter(n != 1) |>  # 86,264 
  arrange(desc(n))
```


```
visitor_id  date        n
04	        2019-04-20	1584		
04	        2019-04-19	1477		
03	        2019-04-09	1309		
06	        2019-04-20	1261		
04	        2019-04-11	1257		
03	        2019-04-23	1147		
03	        2019-04-26	1092		
03	        2019-04-20	1069		
06	        2019-04-19	1004		
03	        2019-04-19	993	
```
That is a bit better, but still rather strange. I worry this short id could be some error? or else? 
Are some tickets sold for more than one day? (its not important for modelling the pathways, but to at least know why are the numbers so high) 

```{r}
visitors_100 <- usage |> count(visitor_id, date) |> #93,595 
  filter(n != 1 & n <= 100) |>  # 86,264 
  arrange(desc(n)) |> select(visitor_id, date) 
```


# Convert single visitor data into network data with origin and destination  

```{r}
visitor <- usage |> filter(visitor_id == "02000000045e147957") # example visitor
```

```{r}
networking <- function(visitor){
  origin <- c()
  dest <- c()
  time_origin <- c()
  time_dest <- c()

  n <- visitor |> nrow()
  
  for(i in 1:(n-1)){
    origin[i] <- visitor$point_id[i]
    dest[i] <- visitor$point_id[i+1]
    # visitor_id[i] <- visitor$visitor_id[i] 
    time_origin[i] <- visitor$timestamp[i] 
    time_dest[i] <- visitor$timestamp[i+1] 
  }
  
  return(data.frame(
    visitor_id = (rep(visitor$visitor_id |> unique(),n = n-1)),
    origin = origin, 
    dest = dest, 
    time_origin = time_origin,  
    time_dest = time_dest, 
    duration = time_dest - time_origin, #time in second in between the two events 
    weekday = rep(visitor$weekday |> unique(),n = n-1) 
  ) |> mutate(time_origin = as.POSIXct(time_origin), 
               time_dest = as.POSIXct(time_dest)))
}
```

```{r}
networking(visitor)
```

Now it needs to be just done for all the visitors - a long computation, so, are we doing it? 

